Hydras \& Co.

### 2024-01-15 doc/hydra-chapter.tex
Improve doc (#172)
* small corrections

* corrections to the general introduction

* corrections to the general introduction

* corrections chapter 2

Moreover, we note that this description is in \emph{imperative} terms. In order to formally study the properties of hydra battles, we prefer to use a mathematical vocabulary, i.e., graphs, relations, functions, etc.
Thus, the replication process will be represented as a binary relation on a data type \texttt{Hydra},
linking the state of the hydra \emph{before} and \emph{after} the transformation.
A battle will thus be represented as a sequence of terms of type \texttt{Hydra}, respecting the rules of the game.

@@

\begin{figure}[hp]
\centering
\begin{tikzpicture}[very thick, scale=0.5]

@@

\begin{figure}[hp]
\centering
\begin{tikzpicture}[very thick, scale=0.6]

@@

\begin{figure}[hp]
\centering
\begin{tikzpicture}[very thick, scale=0.4]

@@

Using this representation, one can re-define all the constructions of this chapter, which is left as an exercise.
You will probably have to use patterns described for instance in~\cite{BC04} or the archives of the \coq communication channels (please consult~\url{https://coq.inria.fr/community.html}).

\end{remark}

@@

We notice that this sub-goal does not contain any hypothesis
on the height and size of the hydra \texttt{h}. So, it looks hard to prove the conclusion. Let's stop.

@@

In this section, we represent the rules of hydra battles as a binary relation associated with
a \emph{round}, i.e., an interaction composed of the two following actions:
\begin{enumerate}
\item Hercules chops off one head of the hydra.
\item Then, the  hydra replicates the wounded part (if the head is at distance $\geq 2$ from the foot).
\end{enumerate}
The relation associated with each round of the battle is parameterized  by the \emph{expected} replication  factor (irrelevant if the chopped head is at distance 1 from the foot,
but present for consistency's sake).

@@

If Hercules chops off a head close to the root, there is no replication at all. We use an auxiliary 
predicate \texttt{S0}, associated with the removing of one head from a sequence of hydras.

@@

Let us represent in \coq{}   the transformation of the hydra of Fig.~\vref{fig:Hy} into
the configuration represented in Fig.~\ref{fig:Hy-prime}.

@@

Let us merge \texttt{R1} and \texttt{R1}  into a single relation.
First,  we define the  relation \texttt{(round\_n n h h')} where \texttt{n} is the expected number of  replications (irrelevant in the case of an \texttt{R1}-transformation).
Then, we define a \emph{round} (small step) of a battle
by abstraction over \texttt{n},

@@

\begin{exercise}
  Prove that if \texttt{$h$ -+-> $h'$}, then
  the height of $h'$ is less or equal than the height of $h$.

@@

@@

\subsection{Testing  \dots}
\label{sect:testing}
In order to study \emph{experimentally} the different  configurations of the  battle, we will use a simple data type for representing the states as tuples composed of
the round number, and the respective number of daughters  \texttt{h2}, \texttt{h1}, and heads
of the current hydra.

@@

The battle we are studying looks to be awfully long. Let us concentrate our
tests on some particular events : the states where $\texttt{nh}=0$.
From the value of \texttt{test 5},  it is obvious that at the 10-th round, the counter \texttt{nh} is equal to zero.

@@

Let us define a binary relation associated with every round of the battle.
In the following definition \texttt{i} is associated with the round number (or date, if we consider a discrete time), and \texttt{a}, \texttt{b}, \texttt{c} respectively associated with the number of occurrences of \texttt{h2}, \texttt{h1} and heads connected to the hydra's foot. For convenience, we do not use the type \texttt{state} of the preceding section, but consider the round numbers and the number of hydras \texttt{h2}, \texttt{h1} and heads as separate arguments of the relation (which is no more ---formally--- ``binary'').

@@

The following lemma establishes a relation between \texttt{steps} and the predicate \texttt{battle}.

@@

The number $N$ is greater than or  equal to $2^{2^{95}\times 95}.$ If we wrote $N$ in base $10$, $N$ would require at least $10^{30}$ digits!

@@

The best known well-founded order is the natural order on the set $\mathbb{N}$ of natural numbers (the type \texttt{nat} of Standard library). It would be interesting to look for some measure $m:\texttt{nat}\arrow\texttt{nat}$ and prove it is a variant.

>>>>

Moreover, we note that this description is in \emph{imperative} terms. In order to formally study the properties of hydra battles, we prefer to use a mathematical vocabulary, i.e., graphs, relations, functions, etc.
Thus, the replication process will be represented as a binary relation on a data type \texttt{Hydra},
linking the state of the hydra \emph{before} and \emph{after} the transformation.
A battle will thus be represented as a sequence of terms of type \texttt{Hydra}, respecting the rules of the game.
 In other terms, we consider hydra battles as \emph{transition systems}.

@@

\begin{figure}[h]
\centering
\begin{tikzpicture}[very thick, scale=0.5]

@@

@@

\begin{figure}[h]
\centering
\begin{tikzpicture}[very thick, scale=0.4]

@@

Using this representation, one can re-define all the constructions of this chapter, which is left as an exercise.
You will probably have to use patterns described for instance in~\cite{BC04, BC04ch14} or the archives of the \coq communication channels (please consult~\url{https://coq.inria.fr/community.html}).
\end{remark}

@@

We notice immediately that the context of this sub-goal does not allow to infer its conclusion.
 Let's stop.

@@

In this section, we represent the rules of hydra battles as a binary relation associated with
a \emph{round}\footnote{usually called a \emph{small step semantics}}, i.e., an interaction composed of the two following actions:
\begin{enumerate}
\item Hercules chops off one head of the hydra.
\item Then, the  hydra replicates the wounded part (if the head is at distance $\geq 2$ from the foot).
\end{enumerate}
The relation associated with each round of the battle is parameterized  by the \emph{expected} replication  factor (irrelevant if the chopped head is at distance 1 from the foot,
but present for consistency's sake).

@@

If Hercules chops off a head next to the root, there is no replication at all. We use an auxiliary 
predicate \texttt{S0}, associated with the removing of one head from a sequence of hydras.

@@

Let us represent in \coq{}   the transformation of the hydra of Fig.~\vref{fig:Hy} into
the configuration represented in Fig.~\vref{fig:Hy-prime}.

@@

Let us merge \texttt{R1} and \texttt{R2}  into a single relation.
First,  we define the  relation \texttt{(round\_n n h h')} where \texttt{n} is the expected number of  replications (irrelevant in the case of an \texttt{R1}-transformation).
Then, we define a \emph{round} (small step) of a battle
by abstraction over \texttt{n},

@@

@@

\index{hydras}{Exercises}
\begin{exercise}
  Prove that if \texttt{$h$ -+-> $h'$}, then
  the height of $h'$ is less or equal than the height of $h$.

\end{exercise}
\index{hydras}{Exercises}

@@

\subsection{Testing  \dots}
\label{sect:testing}
In order to make the study of this battle easier,  we will use a simple data type for representing a configuration 
$(\textit{round}, \texttt{hyd}\;n_2\;n_1\; n_h)$ as the $4$-tuple

@@

The battle we are studying looks to be awfully long. Let us concentrate our
tests on some particular events : the states where $n_h=0$.
From the value of \texttt{test 5},  it is obvious that at the 10-th round, the counter \texttt{nh} is equal to zero.

@@

Let us define a binary relation associated with every round of the battle.
In the following definition \texttt{i} is associated with the round number (or date, if we consider a discrete time), and \texttt{a}, \texttt{b}, \texttt{c} respectively associated with the number of occurrences of \texttt{h2}, \texttt{h1} and heads connected to the hydra's foot. For convenience\footnote{In a few words, the type \texttt{state} was designed for performing \emph{computations}, and \texttt{steps} for writing \emph{interactive proofs}, inspired by the aforementionned computations.},
we do not use the type \texttt{state} of the preceding section, but consider the round numbers and the number of hydras \texttt{h2}, \texttt{h1} and heads as separate arguments of the relation (which is no more ---formally--- ``binary'').

@@

The following lemma establishes a relation between \texttt{steps} and the predicate \texttt{rounds}.

@@

The number $N$ is greater than or  equal to $2^{2^{95}\times 95}.$ If we write $N$ in base $10$, $N$ would require at least $10^{30}$ digits!

@@

The best known well-founded order is the natural order on the set $\mathbb{N}$ of natural numbers (the type \texttt{nat} of Standard Library). It would be interesting to look for some measure $m:\texttt{nat}\arrow\texttt{nat}$ and prove it is a variant.

### 2024-01-15 doc/schutte-chapter.tex
Fix a typo
Update doc/schutte-chapter.tex
Co-authored-by: Columbus240 <8899730+Columbus240@users.noreply.github.com>
Update doc (#161)
* fix a few Hint locality attributes

* Document the use of zorns-lemma

* Fix a typo

* Update doc/thebib.bib

Co-authored-by: Columbus240 <8899730+Columbus240@users.noreply.github.com>

* Update theories/ordinals/Schutte/Countable.v

Co-authored-by: Columbus240 <8899730+Columbus240@users.noreply.github.com>

* Update doc/schutte-chapter.tex

Co-authored-by: Columbus240 <8899730+Columbus240@users.noreply.github.com>

* Update theories/ordinals/Schutte/Countabl

Countable sets were

@@

Is it true ?

\emph{You may start this exercise with the file
    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/schutte_cnf_counter_example.v}{exercises/ordinals/schutte\_cnf\_counter\_example.v}.}
\end{exercise}

@@

\begin{remark}
The sub-directory
    \href{https://github.com/coq-community/hydra-battles/tree/master/theories/ordinals/Gamma0}{theories/oridinals/Gamma0} contains an (incomplete, still undocumented) implementation of the set of ordinals below $\Gamma_0$, represented in Veblen normal form. 
\end{remark}

>>>>

@@

Is it true ?
\emph{You may start this exercise with the file
    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/schutte_cnf_counter_example.v}{exercises/ordinals/schutte\_cnf\_counter\_example.v}.}
\end{exercise}

@@

\begin{remark}
The sub-directory
    \href{https://github.com/coq-community/hydra-battles/tree/master/theories/ordinals/Gamma0}{theories/ordinals/Gamma0} contains an (incomplete, still undocumented) implementation of the set of ordinals below $\Gamma_0$, represented in Veblen normal form. 
\end{remark}
